\documentclass[a4paper,12pt]{article}

% ==========================================
% 1. CONFIGURACIÓN DE IDIOMA Y CODIFICACIÓN
% ==========================================
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel} % Usamos inglés para el paper de arXiv. Si lo quieres en español, cambia a 'spanish'

% ==========================================
% 2. PAQUETES MATEMÁTICOS Y FÍSICA AVANZADA
% ==========================================
\usepackage{amsmath, amssymb, amsthm, mathtools} % Matemáticas y Teoremas
\usepackage{physics}                             % Notación bra-ket, derivadas, etc.
\usepackage{tensor}                              % Índices tensoriales (crucial para Riemann-Cartan)
\usepackage{bm}                                  % Vectores y tensores en negrita
\usepackage{slashed}                             % Para ecuaciones de Dirac (si las usamos)

% ==========================================
% 2.1 TIPOGRAFÍA 
% ==========================================
\usepackage{microtype} % <--- Mejora el espaciado y la lectura.

% ==========================================
% 3. ENTORNOS DE TEOREMAS (¡CRUCIAL!)
% ==========================================
% Esto es lo que faltaba en tu propuesta para que funcionen las Secciones 2 a 5
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{axiom}{Axiom}[section]
\newtheorem{postulate}{Postulate}[section]

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

% ==========================================
% 4. GRÁFICOS, CIRCUITOS Y DISEÑO
% ==========================================
\usepackage{graphicx}
\usepackage{float}
\usepackage{circuitikz} % ¡Sí! Lo añadimos para el Transistor Cuántico v3.2
\usepackage{geometry}
\geometry{margin=2.5cm}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

%% ==========================================
% 5. COMANDOS PERSONALIZADOS GEM
% ==========================================
% Comandos generales
\newcommand{\LGEM}{\mathcal{L}_{\text{GEM}}}
\newcommand{\alphaGEM}{\alpha_{\text{topo}}}

% Comandos de la Sección 2 (Geometría Riemann-Cartan)
\newcommand{\torsion}{T}
\newcommand{\vierbein}{e}
\newcommand{\curvature}{R}
\newcommand{\manifold}{\mathcal{M}}
\newcommand{\connection}{\Gamma}
\newcommand{\contortion}{K}
\newcommand{\spinsconn}{\omega}
\newcommand{\lagrangian}{\mathcal{L}}
\newcommand{\action}{\mathcal{S}}
\newcommand{\vacuum}{\mathcal{V}}
\newcommand{\scalarfield}{\varphi}
\newcommand{\fibbundle}{\mathcal{P}}
\newcommand{\structgrp}{G}
\newcommand{\baseM}{M}

% ==========================================
% INICIO DEL DOCUMENTO
% ==========================================
\begin{document}

% Metadatos del PDF
\hypersetup{
    pdftitle={Topological Vacuum Structure and the Geometric Origin of Fundamental Couplings},
    pdfauthor={Ivan Ugidos Martinez},
    pdfsubject={Riemann-Cartan geometry, icosahedral vacuum},
    pdfkeywords={torsion, Nieh-Yan invariant, topological defects},
    pdfproducer={GEM Research - Version 1.1}
}

\title{\textbf{Topological Vacuum Structure and the Geometric Origin of Fundamental Couplings: A Riemann-Cartan Approach}\thanks{Version 1.1 --- Revised manuscript addressing reviewer feedback on dimensional analysis, symmetry breaking scales, and mass hierarchy predictions.}}
\author{Iván Ugidos Martinez \\ \small \texttt{espejogalactico@gmail.com}}
\date{\today}

\maketitle
\vspace{1cm}
\noindent \textbf{Copyright and License:} \\
\noindent \textcopyright \space 2026. This work is licensed under a \href{https://creativecommons.org/licenses/by-sa/4.0/}{Creative Commons Attribution-ShareAlike 4.0 International License (CC BY-SA 4.0)}. \\
\noindent You are free to share and adapt this material for non-commercial scientific research, provided appropriate credit is given.

\begin{abstract}
The Standard Model of particle physics successfully describes fundamental interactions but leaves the origin of its dimensionless parameters as empirical inputs. In this paper, we propose a rigorous topological reformulation of the quantum vacuum, modeling it as a Riemann-Cartan manifold endowed with non-vanishing torsion and a spontaneously broken discrete icosahedral substructure ($I_h \subset SO(3) \subset SO(1,3)$) at the Planck scale. By extending Emmy Noether’s theorems to this torsioned geometry, we provide a topological estimation of the fine-structure constant $\alpha$ and show that electric charge quantization emerges naturally from the fundamental group of the icosahedral order parameter space. Furthermore, we identify a topological correction to the muon anomalous magnetic moment arising from the discrete vacuum structure. However, we explicitly recognize a hierarchy problem of approximately 34 orders of magnitude between the fundamental Planck scale and the effective electroweak coupling scale, leaving its resolution as an open problem analogous to the Higgs hierarchy. We conclude by proposing falsifiable predictions for high-energy astrophysics and an exploratory hypothesis for macroscopic negentropic cooling.
\end{abstract}

\newpage
% Opción A: Solo mostrar Secciones principales (1, 2, 3... 7)
%\setcounter{tocdepth}{1} 
% Opción B: Mostrar hasta Subsecciones (2.1, 2.2... 7.3), pero ocultar 7.3.1
\setcounter{tocdepth}{2}

\tableofcontents
\newpage

% ==========================================
% SECCIONES DEL DOCUMENTO (Modularidad)
% ==========================================

\input{01_Introduction}

\input{02_Mathematical_Framework}

\input{03_Derivation_Alpha.tex}

\input{04_Charge_Quantization.tex}

\input{05_Fermion_Generations}

\input{06_Conclusions}

\input{07_Open_Problems}


%\input{07_GEM_Teorema_EM}

% ==========================================
% APÉNDICE: EL TRANSISTOR CUÁNTICO (Circuitikz)
% ==========================================
%\appendix
%\section{Hardware Implementation: The GEM-Lite v3.2 Transistor}
%\input{Appendix_A_Transistor_Circuit}

% ==========================================
% BIBLIOGRAFÍA
% ==========================================
\begin{thebibliography}{9}

\bibitem{Cartan1922}
E. Cartan, \textit{On a generalization of the notion of Riemannian curvature and the notion of torsion}, 
C. R. Acad. Sci. Paris \textbf{174}, 593 (1922).

\bibitem{Hehl1976}
F. W. Hehl, P. von der Heyde, G. D. Kerlick, J. M. Nester, \textit{Spin and torsion in gravitation}, 
Phys. Rev. D \textbf{12}, 3139 (1975).

\bibitem{NiehYan1982}
H. T. Nieh and M. L. Yan, \textit{An identity in Riemann-Cartan geometry}, 
J. Math. Phys. \textbf{23}, 373 (1982).

\bibitem{Shapiro2002}
I. L. Shapiro, \textit{Physical Aspects of the Space-Time Torsion}, 
Phys. Rept. \textbf{357}, 113 (2002).

\bibitem{Mielke2017}
E. W. Mielke, \textit{Geometrodynamics of Gauge Fields}, 
Springer (2017).

\end{thebibliography}

\end{document} 